%!TEX root = ../dissertation.tex
% the acknowledgments section

This thesis arose from group projects in two classes, and I am forever indebted
to my project partners for going on this wild ride with me.
To Emily Wang, who took me seriously when I suggested doing a project on
flocking more than a year ago, and to Thomas Lively, who jumped onto this weird
project when I wanted to try out a new direction.
The quality of this work has been immeasurably improved by my collaboration
with the two of you, and there is no doubt in my mind that this work would not
have come to fruition without your help.

I am deeply grateful for the two advisors I've had the pleasure of working with
over the course of this thesis, Barbara Grosz and Radhika Nagpal
Their comments have sharpened this work over the past year and helped shape it
into the form it takes today.
Special thanks are due to Radhika, who agreed in December to be my former
thesis advisor with just four months' notice.
My other thesis reader and esteemed professor Yiling Chen also deserves my
heartfelt thanks.
And last but certainly not least, the dear friends of mine who read and
commented on early drafts of this document: Josie Lo Bello, who read
my thesis in the middle of the night on a thirteen-hour flight to China; Wendy
Woodin, who gave me detailed comments on her lunch break because I asked the
day before this thesis was due; Ross Rheingans-Yoo, who was a great peer
mentor when I was an underclassman and who has continued to be a great friend
from across the world in Hong Kong; and my brother Matthew Fu, who has been
supporting me in countless ways since I was four.

I would be remiss not to mention the many professors, chief among them Margo
Seltzer and Eddie Kohler, who did not directly contribute to this project but
who have nevertheless shaped my undergraduate experience in innumerable ways.
Over the years, I have TF'd CS 61 with Margo, Eddie, and both Margo and Eddie
at the same time, and there's nothing quite like seeing three different
iterations of the same course to help you appreciate your mentors.
I am especially thankful for the many ways Margo has helped me along my
research journey over the years, from pointing me to a research-based startup
when I e-mailed her looking for a summer internship in June of my freshman year
to taking me onto the ASC project in my junior year.

My deepest gratitude goes out to the Harvard Ballroom Dance Team, which has
helped me grow in so many ways over the past four years and among whose team
members I can count my closest friends.
Shout outs as well to my friends in MD 121, who always kept their doors open
for me and gave me comfort in times of need.

And finally, I'd like to thank my parents Henry Fu and Hillary Wu for always
believing me and supporting me over the years.
Without their love and support, I would not be where I am today. 
